---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data


```{r, echo=TRUE}

unzip("D:/Users/philcarmine/Documents/RepData_PeerAssessment1-master/RepData_PeerAssessment1-master/activity.zip")
dataset <- read.csv("D:/Users/philcarmine/Documents/RepData_PeerAssessment1-master/RepData_PeerAssessment1-master/activity.csv")
library("dplyr")
totalperday <- aggregate(dataset$steps, by=list(dataset$date), FUN=sum)
totalperday <- data.frame(Date=totalperday$Group.1, Steps=totalperday$x)

```

## Histogram of total number of steps taken each day

```{r totalperday=TRUE}
steps <- totalperday$Steps
steps <- na.omit(steps)
hist(steps)

```

## What is mean and median total number of steps taken per day?

```{r steps, echo = TRUE}

data.frame(Mean = mean(steps), Median = median(steps))


```

## Time series plot showing average steps per day

```{r dataset}
datasetNA <- na.omit(dataset)
avgstepsperday <- aggregate(datasetNA$steps, by=list(datasetNA$date), FUN=mean)
avgstepsperday <- data.frame(Date = avgstepsperday$Group.1, AvgStepsPerInterval = avgstepsperday$x)
with(avgstepsperday, (plot(avgstepsperday$Date, avgstepsperday$AvgStepsPerInterval, type = "n", xlab = "Date", ylab = "AvgStepsPerInterval")))
points(avgstepsperday$Date, avgstepsperday$AvgStepsPerInterval, type = "l")

```

## 5-minute Interval with highest average  

```{r datasetNA}
intervalstepavg <- aggregate(datasetNA$steps, by=list(datasetNA$interval), FUN = mean)
intervalstepavg <- data.frame(Interval = intervalstepavg$Group.1, AvgSteps = intervalstepavg$x)
intervalstepavg <- intervalstepavg[order(intervalstepavg$AvgSteps),]
tail(intervalstepavg, 1)

```

## Imputing missing values

As there are entire days with missing values, the strategy taken to impute missing values will be to impute the mean number of steps for that interval across all days into the interval for any given day that the value is missing

```{r}

datasetimputed <- inner_join(dataset, intervalstepavg, by = c("interval" = "Interval"))
datasetimputed$steps <- ifelse(is.na(datasetimputed$steps), datasetimputed$AvgSteps, datasetimputed$steps)

```

## Total steps per day (after imputing interval mean for missing values)

```{r datasetimputed}

totalperdayimputed <- aggregate(datasetimputed$steps, by=list(datasetimputed$date), FUN=sum)
hist(totalperdayimputed$x, xlab = "Total Steps per Day", main = "Histogram of Total Steps per Day", breaks = 12)

```

## Are there differences in activity patterns between weekdays and weekends?

```{r}

library(chron)
isweekend <- is.weekend(datasetimputed$date)
isweekend <- data.frame(isweekend = isweekend)
datasetimputed <- cbind(datasetimputed, isweekend)
weekendanalysis <- aggregate(datasetimputed$steps, by=list(datasetimputed$interval, datasetimputed$isweekend), FUN=mean)
par(mfrow = c(2,1))
with(weekendanalysis, plot(Group.1, x, type = "n", xlab = "Interval", ylab = "Number of Steps", main = "Weekend"))
  with(subset(weekendanalysis, Group.2 == "TRUE"), points(Group.1, x, type = "l"))
  with(weekendanalysis, plot(Group.1, x, type = "n", xlab = "Interval", ylab = "Number of Steps", main = "Weekday"))
  with(subset(weekendanalysis, Group.2 == "FALSE"), points(Group.1, x, type = "l"))

```

